GEE models

### 2017-02-22 doc/submission/reviewer-response.md

# Responses to reviewers

## Referee: 1

- *"First, I would like to congratulate you for the exhaustive and complete
research work that has been done for this paper. I think the statistical study
is very successful and useful for researchers who have accumulated a large
database and need a method with the ability to handle a large number of
variables and their interactions. I really liked read in the discussion the
negative points of the study, with which I fully agree, and not an attempt to
hide the limitations of any scientific work. In retrospect, perhaps it will miss
data that **clarify the pre-diabetic state, or not, of the patients (HbA1c or
C-peptide)**. I think it would be good for the study know if some patients were
taking **any medication that could affect the results**. Although it is a
complicated scientific work to understand for researchers who do not usually use
this kind of statistical studies, I believe that is described in a very
appropriate and clear manner."*

    We thank the reviewer for these comments. {{not sure about comment on
    pre-diabetes}}. In the PROMISE cohort, we collected medication data from
    all participants. While most participants took some form of medication over
    the 6 years, during the times of data collection, few ({{get number}}) were
    on medications. However, for those who were taking medications that could
    influence the results, sensitivity analyses revealed they did not
    substantially influence NEFA concentrations, the outcome measures, nor the
    GEE model results {{need to run/confirm this}}. These results have been
    appended to the ESM {{yes?}}.

@@

1. *"Title of the paper: it is a statement of the results. I would prefer to leave
it open ended without an interpretation of the results, as with different levels
of adjustments the results and conclusion will change substantially. See also
comments on the results section."*

    As suggested, we've made the title results agnostic.

2. *"Methods in the abstract: a) Please give some information on the cohort
characteristics: age, sex, high risk of diabetes but no prevalent disease at
baseline; b) incident diabetes cases (how were these handled? what about treated
patients? - these should go in the main methods section); c) adjustment for time
– I would prefer to state that time since baseline was used as the underlying
time variable instead of adjusting for time"*

    We have made the requested changes to the methods and stated the exact
    analyses used more explicitly. Incident diabetes cases are presented only
    for descriptive purposes and were not included in any further analysis. We
    thank the reviewer for the suggestion and agree that using time since
    baseline is a better variable to adjust for; the time variable has been
    replaced in GEE models, though the change does not substantially impact the
    final results. {{confirm}}

@@

6. *"MATERIALS & METHODS:"*
    - *"ESM Figure 1 is not clear to me: a) Prevalent diabetes cases were excluded
    at baseline (n=XX); b) the lines suggest that everyone had to have a year 6
    visit, ehile in the main text, the authors state that any follow-up visit
    was enough for being included (this corresponds well with the GEE
    requirements); c) what happened with incident diabetes cases? Were treated
    diabetes cases included in the analysis? How does that effect measures of
    insulin sensitivity and beta-cell function? I suggest to exclude those
    visits where diabetes was treated."*

        We have clarified how we dealt with incident diabetes cases. Briefly, we
        only report incident diabetes and dysglycemia cases in the text, but in
        the modeling these were excluded. {{not sure how to address a) and b)...}}

    - *"I would prefer to see the equations for the outcome measures. Why not use
    HOMA2-IS instead of 1/HOMA-IR?"*

        We have included the equations for the outcome measures as requested. In
        addition, we replaced HOMA-IR with HOMA2-%S from the HOMA2 Calculator.
        The results were not substantially different after the replacement of
        HOMA-IR. {{confirm}}

@@

7. *"RESULTS:"*
    - *"Please give ns in Table 1."* ... *"It would be also nice to see, how
    many participants had 1, 2 or 3 time points in the analysis."*

        These have been added as requested.

    - *"How were the changes calculated (GEE?) in the basic characteristics of
    the PROMISE cohort section?"*

        {{not completely sure what is meant here}} The changes over time as
        reported in the results of the outcome variables are simple median
        differences between the baseline and 6-year visit. The p-value {{confirm
        this is added}} representing the difference was computed using an
        unadjusted GEE model.

@@

    - *"I suggest to include the unadjusted data to the main paper (ESM Figure 3)
    and describe these findings as we do not know the exact model where
    over-adjustment start to happen."*

        The unadjusted results and figure has been added to the main paper and
        described in the results section. We've also added a sensitivity
        analysis confirming which covariate(s) attenuates the associations.

@@

    - *"In a trajectory analysis, the interaction between time and the main
    predictor (NEFA in this case) is of major interest. The authors check for
    this interaction only as the last step of their model building (see ESM). I
    would suggest to perform this as the first step of the analysis (M0), and
    drop this term, if it is non-significant or does not improve model fit.
    Furthermore, given that this interaction was dropped from the model, it is
    hard to except the major conclusion that NEFA predicts progression of
    beat-cell dysfunction. Actually what is found here is that NEFA are related
    to beta-cell function in repeated cross-sectional analyses and NEFA related
    differences in beta-cell function remain constant over time."*

        The reviewer makes an excellent point regarding the interaction check in
        model building. We have incorporated this change in the analysis and
        report on the results in the paper. In simple terms, the reviewer is
        correct that without an interaction term the model is essentially
        "repeated cross-sectional analyses", however, GEE handles the analysis
        in a slightly more nuanced approach. GEE calculates the estimates of the
        association of NEFA on the outcomes at any given time point, taking into
        consideration the inherent multivariate structure of the repeated
        measurements of the outcome variables. As such, inherent similarities in
        the outcome measurements over time within a subject are included in the
        computation of the GEE estimates. This nuance provides substantially
        more power to the model and the proceeding results.

@@

        Given this reviewer's previous comments and suggestions, we have decided
        to focus on the GEE analysis in the main paper and move the PLS-DA to
        the ESM and expand on {{or remove? it seems this reviewer prefers if we
        remove it. I could simplify it by doing what I did in the TAG as a
        confirmation, and include it in the ESM}} it there.

    - *"The authors mention in the Discussion the external validity is mostly
    limited to white females. I suggest to run a sensitivity analysis on this
    group."*

        We agree with the reviewer's suggestion and ran additional sensitivity
        analyses stratified by sex and ethnicity. We mention the results of this
        analysis in the results.

8. *"DISCUSSION"*
    - *"I would like to see more on the DAGs that underlies the selected
    variables in the models. That would either confirm the way the analysis is
    currently performed, or would show that some of the adjustments are
    over-adjustment."*

### 2017-06-11 doc/submission/reviewer-response.md

---
output: fost::default_manuscript
---

# Responses to reviewers

## Referee: 1

- *"First, I would like to congratulate you for the exhaustive and complete
research work that has been done for this paper. I think the statistical study
is very successful and useful for researchers who have accumulated a large
database and need a method with the ability to handle a large number of
variables and their interactions. I really liked read in the discussion the
negative points of the study, with which I fully agree, and not an attempt to
hide the limitations of any scientific work. In retrospect, perhaps it will miss
data that clarify the pre-diabetic state, or not, of the patients (HbA1c or
C-peptide). I think it would be good for the study know if some patients were
taking any medication that could affect the results. Although it is a
complicated scientific work to understand for researchers who do not usually use
this kind of statistical studies, I believe that is described in a very
appropriate and clear manner."*

### 2017-07-01 doc/manuscript.Rmd
Cleaned up first setup chunks manuscript

```{r knit_setup, echo=FALSE}
knitr::opts_chunk$set(echo = FALSE, dev = c('cairo_ps'))
```

```{r setup, message=FALSE, echo=FALSE}
devtools::load_all('.')
ds <- project_data
ds0yr <- dplyr::filter(ds, VN == 0)
```

```{r caption_setup}
## Include captions below using `captioner` package
library(captioner)
fig <- captioner(prefix = 'Fig.')
cite_f <- pryr::partial(fig, display = 'cite')
sfig <- captioner(prefix = 'ESM Fig.')
cite_sf <- pryr::partial(sfig, display = 'cite')
tab <- captioner(prefix = 'Table')
cite_t <- pryr::partial(tab, display = 'cite')
stab <- captioner(prefix = 'ESM Table')
cite_st <- pryr::partial(stab, display = 'cite')
# usage: cite_st('basicChar')

>>>>

```{r setup, echo=FALSE}
devtools::load_all('.')
```

```{r captions}
## Include captions below using `captioner` package
library(captioner)
fig <- captioner(prefix = 'Fig.')
cite_f <- pryr::partial(fig, display = 'cite')
sfig <- captioner(prefix = 'ESM Fig.')
cite_sf <- pryr::partial(sfig, display = 'cite')
tab <- captioner(prefix = 'Table')
cite_t <- pryr::partial(tab, display = 'cite')
stab <- captioner(prefix = 'ESM Table')
cite_st <- pryr::partial(stab, display = 'cite')
# usage: cite_st('basicChar')
